---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SapCal

<!-- badges: start -->
<!-- badges: end -->

The goal of SapCal is to calculate tree sap flow using HFD (Nadezhdina et al. 2012) and LHB methods (Trcala and Cermak 2016). This package is created associated with the study "Inconsistent results of sap flow calculations from Heat Field Deformation (HFD) and Linear Heat Balance (LHB) methods" (Zhao et al., to be published).

## Installation

You can install the package of SapCal from [GitHub](https://github.com/)  with:


``` r
install.packages("remotes")
remotes::install_github("junbinzhao/SapCal")
```
## Example

1. Calculate the sap flow density (SFD) using the LHB approach
```{r}
library(SapCal)
library(tidyverse)

# load example data
df <- read.csv(file = system.file("extdata", "HFD2J704_202006.csv", package = "SapCal"),
header=T)
names(df)

# convert the data into long format
df <- df %>%
  pivot_longer(cols = Temp1U:Temp8S,
               names_to = c(".value","Position"),
               names_sep = c(5)) %>%
  pivot_longer(cols = starts_with(c("K","Temp")),
               names_to = c(".value","Depth"),
               names_sep = -1) %>%
  pivot_wider(names_from = Position,
              values_from = Temp)

# calculate sap flow density using dynamic K values that are provided as a variable
df_lhb1 <- Cal_LHB(df, # input data
                   T_up = "U", # column name (string) for temperature of the upper sensor
                   T_low = "L", # the lower sensor
                   T_side = "S", # the side sensor
                   K="K", # column name for K values
                   Heat=2.6 # heating power used, W m-1
                   )
tapply(df_lhb1$SFD_lhb,df_lhb1$Depth,summary)

# calculate sap flow density using constant K value
df_lhb2 <- Cal_LHB(df,T_up = "U",T_low = "L",T_side = "S",
                   K=.8, # use a fixed number for the K value
                   Heat=2.6)
tapply(df_lhb2$SFD_lhb,df_lhb2$Depth,summary)
```

2. Calculate the SFD using the HFD approach

```{r}
# calculate sap flow density using dynamic K values that are provided as a variable
df_hfd1 <- Cal_HFD(df,
                   T_up = U, # column name (non-string) for temperature of the upper sensor
                   T_low = L, # the lower sensor
                   T_side = S, # the side sensor
                   K=K, # column name for K values
                   L=7.5 # sapwood depth, cm
                   )
tapply(df_hfd1$SFD_hfd, df_hfd1$Depth, summary)

# calculate sap flow density using constant K value
df <- df %>% select(-K) # remove the K value column to avoid confusion 
df_hfd2 <- Cal_HFD(df,T_up = U,T_low = L,T_side = S,
                   K=.6, # use a fixed number for the K value
                   L=7.5)
tapply(df_hfd2$SFD_hfd, df_hfd2$Depth, summary)
```


